Concurrent, Parallel Garbage Collection in Linear Time

### 2016-06-02 finaletd/paper.tex
algo updated
shared edit
shared

\begin{figure}[!t]
\centering
\includegraphics[width=0.32\textwidth]{figs/simplecyclenew}
\caption{When root R1 is removed, the garbage collection algorithm
only needs to trace link 2 to prove that object A does not need
to be collected.}
\label{prove}
\end{figure}

@@

\begin{figure*}[!t]
  \centering
  %\subfloat[Initially, the owner node $v$ publishes the object]
  {\includegraphics[height=4.5in]{figs/method}\label{fig:example1}}
  %\hspace{30pt}%
  %\subfloat[Node $u$ issues a ${\sf move}$ request]
  %{\includegraphics[height=2.0in]{figs/method1}\label{fig:example2}}\\
  %\hspace{12pt}%
  %\subfloat[The request continues up phase]
  %{\includegraphics[height=2.0in]{figs/method2}\label{fig:example3}}
  %\hspace{30pt}%
  %\subfloat[Directory path found, new owner node $u$]
  %{\includegraphics[height=2.0in]{figs/method3}\label{fig:example4}}
  \caption{Reclaiming a cycle with three objects}%
  \label{ex1}
\end{figure*}

\begin{figure*}[!t]
  \centering
  %\subfloat[Initially, the owner node $v$ publishes the object]
  {\includegraphics[height=2.0in]{figs/doublylinkedlist}\label{fig:example2}}
 \caption{Doubly-linked list}%
  \label{ex2}
\end{figure*}

@@

\begin{figure*}[!t]
  \centering
  %\subfloat[Initially, the owner node $v$ publishes the object]
  {\includegraphics[height=2.0in]{figs/recoveringdbl}}%\label{fig:example1}}

@@

\begin{figure}[t]
\begin{minipage}[t]{0.5\linewidth}
  \centering
% %   %\subfloat[Initially, the owner node $v$ publishes the object]
  \includegraphics[height=1.18in]{figs/graphmodel2}
  %\label{fig:graph1}}
% %   \hspace{12pt}%
  \caption{Graph model}%
\label{fig:graphmodel2}
\end{minipage}
\begin{minipage}[t]{0.5\linewidth}
%\end{wrapfigure}
%\begin{wrapfigure}{r}{1.0in}
  \centering
% %   %\subfloat[Initially, the owner node $v$ publishes the object]
  \includegraphics[height=1.18in]{figs/graphmodel2b}
  %\label{fig:graph1}}
% %   \hspace{12pt}%
  \caption{Subgraph model}%
\label{fig:graphmodel3}
\end{minipage}
\end{figure}

@@

\begin{figure}[h!]
  \centering
% %   %\subfloat[Initially, the owner node $v$ publishes the object]
  \includegraphics[height=1.8in,width=3.0in]{figs/CleanupTimeperObjectonly16log}
  %\label{fig:graph1}}
% %   \hspace{12pt}%
  \caption{
  A large number of independent rings are collected by various number of
  worker threads. Collection speed drops linearly with the number of
  cores used.
  }
   \label{fig:scale}
  \end{figure}

% % While the algorithm can scale very well with the independent sets, the collector threads working on the shared structure needs synchronization. The testcase that we used for scaling is changed so that all benzene structures are interlinked and collector has to communicate and coordinate their activities based on other collector threads. The algorithm systematically achieves optimized communication and coordination among the multiple collector threads working on the same shared structure. As a result of this, the Fig \ref{fig:overhead} shows the synchronization issues due to multiple threads interacting among each other to cleanup the structure.
\begin{figure}[h!]
  \centering
% %   %\subfloat[Initially, the owner node $v$ publishes the object]
  \includegraphics[height=1.8in,width=3.0in]{figs/collapseoverheadonly16log}
  %\label{fig:graph1}}
% %   \hspace{12pt}%
  \caption{A chain of linked cycles is created in memory. The connections are severed,
  then the roots are removed. Multiple collector threads are created and operations
  partially overlap.}
   \label{fig:overhead}
  \end{figure}

\begin{figure}[h!]
  \centering
% %   %\subfloat[Initially, the owner node $v$ publishes the object]
  \includegraphics[height=1.8in,width=3.0in]{figs/linearity}
  %\label{fig:graph1}}
% %   \hspace{12pt}%
  \caption{
  Graphs of different types are created at various sizes in memory,
  including cliques, chains of cycles, large cycles, and large doubly
  linked lists. Regardless of the type of object, collection time
  per object remains constant, verifying the linearity of the underlying
  collection mechanism.
  }
   \label{fig:linearity}
  \end{figure}

@@

\section{Conclusion}
\label{section:conclusion}
We have described a garbage collector based on strong and weak reference counts,
proven its validity,
and illustrated its potential advantages over existing systems, including:
\begin{enumerate}
\item It can run at the same time as a live system, using multiple threads if that
is desired, without needing to ``stop the world.''
%It thus may prove useful in soft real-time systems;
\item It has a reduced need to operate on memory compared to collectors because
it only performs nontrivial work when the last strong link is removed;
\item It has a reduced need to trace links in performing its operations, specifically:
\begin{enumerate}
\item When objects do not need to be collected, it is often able to prove this
without completely tracing cycles to which they belong which have support;
\item It remembers stable paths through memory and thus avoids doing work on old objects,
a benefit similar to that derived from generational collectors, e.g. \cite{Printezis:2000};
deletions are occurring;
\item When objects do need to be collected, the collector only needs to trace the cycle twice;
\item The collector operation is local;
\item The collector does not need or use back-pointers;
\end{enumerate}
These advantages should make our collector useful for distributed systems, where traces that
cross node boundaries are likely to be extremely expensive.
\end{enumerate}

Disadvantages include:
\begin{enumerate}
\item An increased memory overhead: three counters and a pointer are required
for each object;
\item An additional cost of pointer creation/mutation;
\item The lack of a fault tolerance protocol;
\item Little effort has, as yet, been devoted to optimization of any implementation. Further reduction of memory and computational overheads may yet be achieved with variations on this algorithm.
%\item The order of deletion in the $CleanUp$ phase is not, at present, guaranteed. This is a potential problem for systems in which finalizers need to run in a certain order, but a system for mitigating this issue is known to us and will be addressed in a future work.
\end{enumerate}

@@

\section{Appendix}
\subsection{Multi-Threaded Collector}
\label{singlethread}
%\input{algo1}
\clearpage

>>>>

\begin{figure}[!t]
\centering
\includegraphics[width=0.52\textwidth]{figs/simplecyclenew}
\caption{When root R1 is removed, the garbage collection algorithm
only needs to trace link 2 to prove that object A does not need
to be collected.}
\label{prove}
\end{figure}

@@

\begin{figure*}[!t]
  \centering
  %\subfloat[Initially, the owner node $v$ publishes the object]
  {\includegraphics[height=6.5in]{figs/method}\label{fig:example1}}
  %\hspace{30pt}%
  %\subfloat[Node $u$ issues a ${\sf move}$ request]
  %{\includegraphics[height=2.0in]{figs/method1}\label{fig:example2}}\\
  %\hspace{12pt}%
  %\subfloat[The request continues up phase]
  %{\includegraphics[height=2.0in]{figs/method2}\label{fig:example3}}
  %\hspace{30pt}%
  %\subfloat[Directory path found, new owner node $u$]
  %{\includegraphics[height=2.0in]{figs/method3}\label{fig:example4}}
  \caption{Reclaiming a cycle with three objects}%
  \label{ex1}
\end{figure*}

\begin{figure*}[!t]
  \centering
  %\subfloat[Initially, the owner node $v$ publishes the object]
  {\includegraphics[height=3.0in]{figs/doublylinkedlist}\label{fig:example2}}
 \caption{Doubly-linked list}%
  \label{ex2}
\end{figure*}

@@

\begin{figure*}[!t]
  \centering
  %\subfloat[Initially, the owner node $v$ publishes the object]
  {\includegraphics[height=3.0in]{figs/recoveringdbl}}%\label{fig:example1}}

@@

\begin{figure}[t]
\begin{minipage}[t]{0.5\linewidth}
  \centering
% %   %\subfloat[Initially, the owner node $v$ publishes the object]
  \includegraphics[height=2.18in]{figs/graphmodel2}
  %\label{fig:graph1}}
% %   \hspace{12pt}%
  \caption{Graph model}%
\label{fig:graphmodel2}
\end{minipage}
\begin{minipage}[t]{0.5\linewidth}
%\end{wrapfigure}
%\begin{wrapfigure}{r}{1.0in}
  \centering
% %   %\subfloat[Initially, the owner node $v$ publishes the object]
  \includegraphics[height=2.18in]{figs/graphmodel2b}
  %\label{fig:graph1}}
% %   \hspace{12pt}%
  \caption{Subgraph model}%
\label{fig:graphmodel3}
\end{minipage}
\end{figure}

@@

\begin{figure}[h!]
  \centering
% %   %\subfloat[Initially, the owner node $v$ publishes the object]
  \includegraphics[height=3.0in,width=4.0in]{figs/CleanupTimeperObjectonly16log}
  %\label{fig:graph1}}
% %   \hspace{12pt}%
  \caption{
  A large number of independent rings are collected by various number of
  worker threads. Collection speed drops linearly with the number of
  cores used.
  }
   \label{fig:scale}
  \end{figure}

% % While the algorithm can scale very well with the independent sets, the collector threads working on the shared structure needs synchronization. The testcase that we used for scaling is changed so that all benzene structures are interlinked and collector has to communicate and coordinate their activities based on other collector threads. The algorithm systematically achieves optimized communication and coordination among the multiple collector threads working on the same shared structure. As a result of this, the Fig \ref{fig:overhead} shows the synchronization issues due to multiple threads interacting among each other to cleanup the structure.
\begin{figure}[h!]
  \centering
% %   %\subfloat[Initially, the owner node $v$ publishes the object]
  \includegraphics[height=3.0in,width=4.0in]{figs/collapseoverheadonly16log}
  %\label{fig:graph1}}
% %   \hspace{12pt}%
  \caption{A chain of linked cycles is created in memory. The connections are severed,
  then the roots are removed. Multiple collector threads are created and operations
  partially overlap.}
   \label{fig:overhead}
  \end{figure}

\begin{figure}[h!]
  \centering
% %   %\subfloat[Initially, the owner node $v$ publishes the object]
  \includegraphics[height=3.0in,width=4.0in]{figs/linearity}
  %\label{fig:graph1}}
% %   \hspace{12pt}%
  \caption{
  Graphs of different types are created at various sizes in memory,
  including cliques, chains of cycles, large cycles, and large doubly
  linked lists. Regardless of the type of object, collection time
  per object remains constant, verifying the linearity of the underlying
  collection mechanism.
  }
   \label{fig:linearity}
  \end{figure}

@@

\section{Conclusion}
\label{section:conclusion}
We have described a garbage collector based on strong and weak reference counts,
proven its validity,
and illustrated its potential advantages over existing systems, including:
\begin{enumerate}
\item It can run at the same time as a live system, using multiple threads if that
is desired, without needing to ``stop the world.''
%It thus may prove useful in soft real-time systems;
\item It has a reduced need to operate on memory compared to collectors because
it only performs nontrivial work when the last strong link is removed;
\item It has a reduced need to trace links in performing its operations, specifically:
\begin{enumerate}
\item When objects do not need to be collected, it is often able to prove this
without completely tracing cycles to which they belong which have support;
\item It remembers stable paths through memory and thus avoids doing work on old objects,
a benefit similar to that derived from generational collectors, e.g. \cite{Printezis:2000};
deletions are occurring;
\item When objects do need to be collected, the collector only needs to trace the cycle twice;
\item The collector operation is local;
\item The collector does not need or use back-pointers;
\end{enumerate}
These advantages should make our collector useful for distributed systems, where traces that
cross node boundaries are likely to be extremely expensive.
\end{enumerate}
\vspace{0.5in}
Disadvantages include:
\begin{enumerate}
\item An increased memory overhead: three counters and a pointer are required
for each object;
\item An additional cost of pointer creation/mutation;
\item The lack of a fault tolerance protocol;
\item Little effort has, as yet, been devoted to optimization of any implementation. Further reduction of memory and computational overheads may yet be achieved with variations on this algorithm.
%\item The order of deletion in the $CleanUp$ phase is not, at present, guaranteed. This is a potential problem for systems in which finalizers need to run in a certain order, but a system for mitigating this issue is known to us and will be addressed in a future work.
\end{enumerate}

@@

\section{Appendix}
\subsection{Multi-Threaded Collector}
\label{singlethread}
\input{algo1}
\clearpage
